---
title: "pted-week3"
output: html_document
---

#### Task 1 - Segmentation

##load libraries
```{r}
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
```

##read data
```{r}
caro <-read_delim("caro60.csv",",")
```
## Measure the distance from every point to every other point within the temporal window of 6 minutes
```{r}
caro <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),   # distance to pos -3 minutes
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -2 minutes
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -1 minutes
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +1 mintues
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),  # distance to pos +2 minutes
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)  # distance to pos +3 minutes
  )
```

####Task 2 - Specify and apply threshold

##calculate the mean distance of nMinus3, nMinus2, nMinus1, nPlus1, nPlus2, nPlus3 for each row
```{r}
caro <- caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3, nMinus2, nMinus1,nPlus1,nPlus2,nPlus3))
  ) %>%
  ungroup() 
```
##explore the Values of stepMean to choose the treshold 
```{r}
ggplot(data=caro, mapping = aes(x=stepMean))+
  geom_histogram(binwidth = 1)

ggplot(data=caro, mapping = aes(y=stepMean))+
  geom_boxplot()

summary(caro$stepMean)
```
#The mean distance of all stepMean values is  6.951. That's what we use as a threshold.

##Apply threshold to sort stepMean into static or moving
```{r}
caro <- caro %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))

caro_filter <- caro %>%
  filter(!static)
```
####Task 3 - Visualize segmented trajectories
```{r}
ggplot(data = caro, mapping = aes(x= E, y=N) )+
  geom_path()+
  geom_point(aes(colour=static))+
  coord_equal()
```
####Task 4 - Segment-based analysis

##create a unique ID for each segment:
```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
}
```
##assigne IDs to subtrajectories and visualize the segments
```{r}
caro <- caro %>%
  mutate(segment_id = rle_id(static))

caro <- caro %>%
  filter(!static)

ggplot(data = caro, mapping = aes(x= E, y=N) )+
  geom_path(aes(colour=segment_id))+
  geom_point(aes(colour=segment_id))+
  coord_equal()+
  labs(title = "All segments (uncleaned)")+
  theme(title=element_text(size=8),legend.position = "none")
```
##remove all segments n<5 and visualize
```{r}
caro_filter_clean<- caro %>%
  group_by(segment_id)%>%
  mutate(n= n())%>%
  filter(n>=5)

ggplot(data = caro_filter_clean, mapping = aes(x= E, y=N) )+
  geom_path(aes(colour=segment_id))+
  geom_point(aes(colour=segment_id))+
  coord_equal()+
  labs(title = "Long segments (removed <5)")+
  theme(title=element_text(size=8),legend.position = "none")
```
####Task 5 - Similarity measures

##load data
```{r}
pedestrian<-read_delim("pedestrian.csv",",")
```

##look at data structure
```{r}
str(pedestrian)

pedestrian
```
##change ID to factor and Datetime to integer
```{r}
pedestrian <- pedestrian%>%
  mutate(
    trajID_fac=as.factor(TrajID), 
    DateTime_integ=as.integer(DatetimeUTC)
    )
```
##check structure again
```{r}
str(pedestrian)
```
##visualize data
```{r}
ggplot(data=pedestrian, mapping = aes(x=E, y=N,))+
  geom_point(data=dplyr::select(pedestrian,-TrajID),color="grey")+
  geom_point(aes(colour=trajID_fac))+
  facet_wrap(~TrajID)+
  labs(title = "Visual comparison of the 6 trajectories", subtitle= "Each subplot highlights a trajectory")+
  theme_light()+
  theme(legend.position = "none")
```
